---
output_dir: "."
title: "RNotebookExample"
author: "Elin Einarsson"
output: html_document
---

```{r load data, message=FALSE}
# load tidyverse
library(tidyverse)
library(rmarkdown) # used for paged_table function
library(kableExtra) # used for table
library(ggpubr) #used for ggarrange function

#read file into R
pop_data <- read_csv("data/gapminder_data.csv")

#create a table with data from European countries in 2007 showing the countries 
#with the largest life expectancy at the top 
euro_data_tbl <- pop_data %>% 
                    filter(continent == "Europe" & year == 2007) %>% 
                    select(-continent, -year) %>% 
                    arrange(desc(lifeExp)) %>% 
                    rename(Country = country, "Population Size" = pop, 
                      "Life Expectancy" = lifeExp, "GDP" = gdpPercap)

```

The results in euro_data_tbl are displayed in the Table below:

```{r kbl-table}
euro_data_tbl %>% 
          kbl() %>%
          kable_styling(bootstrap_options = c("striped", "hover"))
```

This is a better way how to display longer tables:

```{r paged-table}
paged_table(euro_data_tbl)

```


# Adding images
Create a new dataset euro_data_fig by filtering the pop_data tibble to contain only data from Europe. Draw a plot to display the lifeExp on the y axis and year on the x axis. Use geom_violin() to draw this as a violin plot to show the distrubution of the data across each year and save it in a euro_plot variable.

```{r euro-fig-data}

#keep only European data and change year to factor
euro_data_fig <- pop_data %>% 
                    filter(continent == "Europe")  %>% 
  #year must be factor, makes a category for each year
                    mutate(year=as_factor(year))



```

The life expectancy for Europe over the years is plotted as following:

```{r euro plot, fig.cap= "Life expectancy in Europe over time (years)"}
#Euro plot
euro_plot <- euro_data_fig %>%
                ggplot(mapping=aes(x=year, y=lifeExp)) +
                geom_violin() +
                stat_summary(fun.y = median, geom = "point")

#draw euro plot
euro_plot

```

```{r uk data fig}
uk_data_fig <- pop_data %>% 
                    filter(country == "United Kingdom")  %>% 
  #year must be factor, makes a category for each year
                    mutate(year=as_factor(year))
```

Now we will display more than one image together
```{r uk data image}
uk_plot <- uk_data_fig %>%
                ggplot(mapping=aes(x=year, y=lifeExp)) +
  # UK one country so better to plot as geom_point
                geom_point()

#draw uk plot
uk_plot
```



# Using multiple images as panels 
Using the ggarrange ()

```{r ggarrange}
#put the images you want, ncol= 
ggarrange(euro_plot, 
          uk_plot,
          # number of columns
          ncol = 2, 
          #number of rows, 2 images i.e. 1 
          nrow = 1, 
          #adds label to the panel
          labels = "AUTO")

```




